Variations on a Bayesian Theme: Comparing Bayesian Models of Referential Reasoning

### 2014-03-18 02_RSA.tex
experiment-related revision

To test these intuitions, Frank \& Goodman recruited 745 US
participants via Amazon's Mechanical Turk to take part in a one-shot
referential game, either as the speaker or the listener \cite{Frank}.
The context of the game always consisted of three objects (e.g.,
Figure~\ref{context}).  Each object had three features: color (red,
blue, or green), shape (circle, square, or cloud), and texture (solid,
or polka-dot, or striped).  Every context had one feature fixed for
all objects and the other two features systematically varied.  The
objects were randomly permuted and the target object was randomly
assigned.  Depending on how many objects shared the same value of a
feature with the target, contexts were classified into 7 categories:
1/1, 1/2 (the target's first feature is unique, and there is another
object sharing the same value of the second feature with the target),
1/3, 2/2 with the same object sharing both features with the target
(which means there is another object identical to the target), 2/2
with different objects sharing the features (such as in
Fig.~\ref{context}), 2/3 and 3/3. The speakers were asked to choose a
feature to refer to the target object.\footnote{More accurately,
  participants were ask to \emph{bet} on features and referents. We
  will discuss this design choice when contrasting it with our own
  replication experiment in Section~\ref{sec:experiment}.}  In order
to measure the perceptual salience, some of the listeners were told
that the speaker was using a language that they did not understand and
then they were asked to guess the intended referent.  Other listeners
were told a feature of the target and asked to choose the intended
referent.

The result of the experiment showed that speakers generally followed
Grice's Maxim of Quantity, i.e., they tried to be informative by
choosing the most unique feature of the target, and that listeners
were influenced by both the perceptual salience and the
informativeness of the utterance that the speaker produced. However,
the data, being quantitative in nature, cannot be fully explained by a
classical qualitative Gricean account of scalar implicatures, so that
a major challenge in terms of formal modeling becomes how to quantify
notions such as informativeness and perceptual salience, and how these
should be integrated to yield qualitative predictions.

>>>>

\begin{figure}[h] \label{fig:fg-all}
  \centering \subfigure[1/1]{\label{fg-1-1}
  \includegraphics[scale=0.2]{pics/fg-1-1.jpg}
} \subfigure[1/2]{\label{fg-1-2}
  \includegraphics[scale=0.2]{pics/fg-1-2.jpg}
} \subfigure[1/3]{\label{fg-1-3}
  \includegraphics[scale=0.2]{pics/fg-1-3.jpg}
}

\subfigure[2/2: identical]{\label{fg-2-2-dup}
  \includegraphics[scale=0.2]{pics/fg-2-2-dup.jpg}
} \subfigure[2/2: different]{\label{fg-2-2-crit}
  \includegraphics[scale=0.2]{pics/fg-2-2-crit.jpg}
} \subfigure[2/3]{\label{fg-2-3}
  \includegraphics[scale=0.2]{pics/fg-2-3.jpg}
}

\subfigure[3/3]{\label{fg-3-3}
  \includegraphics[scale=0.2]{pics/fg-3-3.jpg}
}

  \caption{Sample contexts for the 7 conditions used in \cite{Frank}. 
  The targets are indicated by dotted lines. 
  Condition $n/m$ means that the target has one feature shared 
  by $n$ objects and another feature shared by $m$ objects in the context
  (including the target itself).
  For $2/2$, either there is an object identical to the target, or the features are shared by different objects.}
\end{figure}

To test these intuitions, Frank \& Goodman recruited 745 US
participants via Amazon's Mechanical Turk to take part in a one-shot
referential game, either as the speaker or the listener \cite{Frank}.
The context of the game always consisted of three objects (e.g.,
Figure~\ref{context}).  Each object had three features: color (red,
blue, or green), shape (circle, square, or cloud), and texture (solid,
or polka-dot, or striped).  Every context had one feature fixed for
all objects and the other two features systematically varied.  The
objects were randomly permuted and the target object was randomly
assigned.  Depending on how many objects shared the same value of a
feature with the target, contexts were classified into 7 categories
(Fig.~\ref{fig:fg-all}):
1/1, 1/2 (the target's first feature is unique, and there is another
object sharing the same value of the second feature with the target),
1/3, 2/2 with the same object sharing both features with the target
(which means there is another object identical to the target), 2/2
with different objects sharing the features, 2/3 and 3/3.  
The speakers were asked to choose a
feature to refer to the target object.\footnote{More accurately,
  participants were ask to \emph{bet} on features and referents. We
  will discuss this design choice when contrasting it with our 
  replication in Section~\ref{sec:experiment}.}  In order
to measure the perceptual salience, some of the listeners were told
that the speaker was using a language that they did not understand and
then they were asked to guess the intended referent.  Other listeners
were told a feature of the target and asked to choose the intended
referent. 
Note that listeners were presented with pictures without dotted lines 
surrounding the target so they did not know the target.
For each category in each of the speaker, salience and listener conditions, 
there were roughly 50 responses\footnote{As we can see, despite the fairly 
large number of participants recruited, the data in each condition 
are not that many. Also there might be concern that some conditions might 
look less interesting than others. Again we will discuss these issues when 
we introduce our own replication in Section~\ref{sec:experiment}. Here
we note that variety is important in order to test the general validity of 
a pragmatic theory. 
Thus the design used in \cite{Frank} served its purposes well, but may be 
adapted when the goal is different.}.

The result of the experiment showed that speakers generally followed
Grice's Maxim of Quantity, i.e., they tried to be informative by
choosing the most unique feature of the target\footnote{Speakers' preference
was not taken into account in \cite{Frank}, so all features were randomly 
assigned and later lumped together in the analysis.}, and that listeners
were influenced by both the perceptual salience and the
informativeness of the utterance that the speaker produced. However,
the data, being quantitative in nature, cannot be fully explained by a
classical qualitative Gricean account of scalar implicatures, so that
a major challenge in terms of formal modeling becomes how to quantify
notions such as informativeness and perceptual salience, and how these
should be integrated to yield qualitative predictions.

### 2014-03-18 04_experiment.tex
experiment-related revision

There are four major design choices in our own experiment that differ
from the original experiment in \cite{Frank}. First of all, the
original experiment used a \emph{betting paradigm}, i.e., participants
were asked to bet over a range of options and they were instructed
that the amount of money that is bet on an option should reflect their
confidence that the option is correct. Even though the betting
paradigm has the merit of providing us with graded reponses from each
individual participant, the caveat is that it is unclear whether it
measures beliefs or actions. This can lead to confusion when we are to
fit the model predictions to the empirical data without knowing
whether they are directly comparable. In addition, since the betting
paradigm is more or less introspective in nature, it tends to be not
very accurate. Thus we used a \emph{forced choice design} instead,
which clearly measures the action, does not rely on introspection and,
technically speaking, provides for a straightforward likelihood
function to be used in subsequent statistical model comparison.
Secondly, since we decided to investigate the influence of the
speaker's preference as well as the listener's perceptual salience, we
focused on contexts equivalent to Figure \ref{context}, which require
reasoning that is highly reminiscent of Gricean accounts of scalar
implicature calculation, and examined different features separately.
Thirdly, we only included the features of color and shape to ensure
the vocabulary to be common knowledge, excluding the feature of
texture. Finally, we did not use dotted lines surrounding an object as
the way to indicate the target to the speaker, as done in \cite{Frank},
because that might unduly emphasize the feature of shape and thus be a
confound in the experiment. Instead we used a small arrow pointing to
the target.

@@

The results of the speaker and listener conditions of our experiment
are as follows.

>>>>

There are four major design choices in our own experiment that differ
from the original experiment in \cite{Frank}. First of all, the
original experiment used a \emph{betting paradigm}, i.e., participants
were asked to bet over a range of options and they were instructed
that the amount of money that is bet on an option should reflect their
confidence that the option is correct. Even though the betting
paradigm has the merit of providing us with graded reponses from each
individual participant, the caveat is that it is unclear whether it
measures beliefs or actions. This can lead to confusion when we are to
fit the model predictions to the empirical data without knowing
whether they are directly comparable. In addition, since the betting
paradigm is more or less introspective in nature, it tends to be not
very accurate. Thus we used a \emph{forced choice design} instead,
which clearly measures the action, does not rely on introspection and,
technically speaking, provides for a straightforward likelihood
function to be used in subsequent statistical model comparison.
Secondly, since we decided to investigate the influence of the
speaker's preference as well as the listener's perceptual salience, we
focused on contexts equivalent to Figure \ref{context}, which require
reasoning that is highly reminiscent of Gricean accounts of scalar
implicature calculation, and examined different features separately.
Thirdly, we only included the features of color and shape, with binary
values (green/blue and square/circle) for two reasons. On the one hand, 
we wanted to ensure the vocabulary to be common knowledge, so we excluded 
the feature of texture. On the other hand, we excluded feature values red 
and cloud as they are significantly more salient than the others, which 
might override the pragmatic effect. 
Finally, we did not use dotted lines surrounding an object as
the way to indicate the target to the speaker, as done in \cite{Frank},
because that might unduly emphasize the feature of shape and thus be a
confound in the experiment. Instead we used a small arrow pointing to
the target.

@@

The results of the speaker and listener conditions of our experiment
are shown as follows.
